---
title: "2022 - Data Analytics for Immersive Environments - CA4 - RDBMS & Linear Regression
  Project"
author: "Joe O'Regan"
date: "2023-01-11"
output:
  pdf_document: default
  html_document: default
subtitle: "Part 1 - Generate and Populate Database"
---

<!-- style for table rows -->
<style>
  .table-striped tr:nth-child(even) {background: #AAFFAA;}
  .table-hover tbody > tr:hover {background-color: #44FF44}
</style>

<!-- clear console -->
```{r setup, include=FALSE}
cat('\014')
#knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "latex")
```

---

# ER Diagram

<!-- show ER Diagram -->
```{r er image, echo=FALSE}
library("png")
pp <- readPNG("daie_ca4_er_diagram.png")
plot.new()
rasterImage(pp, 0, 0, 1, 1)
```

---

# Create Tables

<!--include libraries-->
```{r setup libraries, include=FALSE}
if(!require("DBI"))
  install.packages("DBI")
if(!require("dplyr"))
  install.packages("dplyr")
if(!require("magick"))
  install.packages("magick")
if(!require("knitr"))
  install.packages("knitr")
if(!require("kableExtra"))
  install.packages("kableExtra")
library(DBI) # interface with database
library(dplyr)
library(magick) # crop images
library(knitr) # Display data in tables
library(kableExtra) # Format tables
```

## Create or recreate the sqlite file

Check if the sqlite file exists already and if it does delete it. Then create an in-memory RSQLite database.

```{r create or recreate the sqlite file, results='hide'}
# if the sqlite file exists already delete it
if (file.exists("daie_ca4_data.sqlite"))
  file.remove("daie_ca4_data.sqlite")

# create the sqlite file
conn <- dbConnect(RSQLite::SQLite(), "daie_ca4_data.sqlite")
```

---

\newpage
## Create the tables

### Asset

Show each table cropped from the main ER Diagram using magick library. Only dimensions and offset change for each cropped table, so this code is shown once.

```{r show asset table only}
er_img <- image_read("daie_ca4_er_diagram.png")
# cropped image dimension offset by location (in pixels)
# (newXdimension x newYdimension + Xlocation + Ylocation)
asset_img = image_scale(image_crop(er_img, "181x117+184+155"), ifelse(is_html_output(), "250%", "125%"))
asset_img
```

Describes a single electronic resource produced by one or more work items

* **Id:** Unique identifier for each row
* **Name:** Description of the asset
* **Type:** Short description of type of asset (Not mandatory)
* **Format:** Short description of format of asset (Not mandatory)
* **Created_By:** Team member who created the asset
* **Date_Created:** Date the asset was created

**Primary Key(s):** Id (auto increments)

**Foreign Key(s):** Created_By (Team_Member.Id)

Create the Asset table. Automatically increment the primary key. Id, Name, Created_By, and Date_created are mandatory fields.

```{r create asset table, results='hide'}
# create tables
dbExecute(conn, "CREATE TABLE Asset (
    Id integer NOT NULL CONSTRAINT Asset_pk PRIMARY KEY AUTOINCREMENT,
    Name varchar(50) NOT NULL,
    Type varchar(50),
    Format varchar(50),
    Created_By integer NOT NULL,
    Date_Created date NOT NULL,
    CONSTRAINT Asset_Team_Member FOREIGN KEY (Created_By)
    REFERENCES Team_Member (Id)
);")
```

---

\newpage
### Asset_Work_Items

```{r show Asset_Work_Items table only, echo=FALSE}
asset_work_items_img = image_scale(image_crop(er_img, "183x62+181+27"), ifelse(is_html_output(), "250%", "125%"))
asset_work_items_img
```

Assets are made up of work items

* **Work_Item_Id:** Id for work items used to create the asset
* **Asset_Id:** Id of the asset

**Primary Key(s):** Composite. Work_Item_Id + Asset_Id

**Foreign Key(s):** Work_Item_Id (Work_Item.Id), Asset_Id (Asset.Id)

```{r create asset_work_items table, results='hide'}
dbExecute(conn, "CREATE TABLE Asset_Work_Items (
    Work_Item_Id integer NOT NULL,
    Asset_Id integer NOT NULL,
    CONSTRAINT Asset_Work_Items_pk PRIMARY KEY (Work_Item_Id,Asset_Id),
    CONSTRAINT Asset_Work_Item_Work_Item FOREIGN KEY (Work_Item_Id)
    REFERENCES Work_Item (Id),
    CONSTRAINT Asset_Work_Items_Asset FOREIGN KEY (Asset_Id)
    REFERENCES Asset (Id)
);")
```

---

\newpage
### Collection

```{r show Collection table only, echo=FALSE}
collection_img = image_scale(image_crop(er_img, "146x61+210+334"), ifelse(is_html_output(), "250%", "125%"))
collection_img
```

Collection of assets

* **Library_Id:** Unique reference for a collection of assets
* **Asset_Id:** Unique reference of the asset that make up a collection

**Primary Key(s):** Composite. Library_Id + Asset_Id

**Foreign Key(s):** Library_Id (Library.Id), Asset_Id (Asset.Id)

```{r create collection table, results='hide'}
dbExecute(conn, "CREATE TABLE Collection (
    Library_Id integer NOT NULL,
    Asset_Id integer NOT NULL,
    CONSTRAINT Collection_Id PRIMARY KEY (Library_Id,Asset_Id),
    CONSTRAINT Collection_Asset FOREIGN KEY (Asset_Id)
    REFERENCES Asset (Id),
    CONSTRAINT Collection_Library FOREIGN KEY (Library_Id)
    REFERENCES Library (Id)
);")
```

---

\newpage
### Library

```{r show Library table only, echo=FALSE}
library_img = image_scale(image_crop(er_img, "136x62+19+331"), ifelse(is_html_output(), "250%", "125%"))
library_img
```

Describes a collection of available assets

* **Id:** Unique reference for each collection in Library
* **Name:** Description of library collection

**Primary Key(s):** Id

**Foreign Key(s):** Id (Collection.Library_Id)

```{r create library table, results='hide'}
dbExecute(conn, "CREATE TABLE Library (
    Id integer NOT NULL CONSTRAINT Library_pk PRIMARY KEY AUTOINCREMENT,
    Name varchar(50) NOT NULL
);")
```

---

\newpage
### Project

```{r show Project table only, echo=FALSE}
project_img = image_scale(image_crop(er_img, "182x74+687+322"), ifelse(is_html_output(), "250%", "125%"))
project_img
```

Describes a project being produced, timeline, and team involved

* **Id:** Unique reference for each project
* **Name:** Description of Project
* **Delivery_Date:** Date project is due to finish

**Primary Key(s):** Id

**Foreign Key(s):** Id (Project_Team.Project_Id). Each Project has multiple Team Members.

```{r create project table, results='hide'}
dbExecute(conn, "CREATE TABLE Project (
    Id integer NOT NULL CONSTRAINT Project_pk PRIMARY KEY,
    Name varchar(128) NOT NULL,
    Delivery_Date date NOT NULL
);")
```

---

\newpage
### Project_Team

```{r show Project_Team table only, echo=FALSE}
project_team_img = image_scale(image_crop(er_img, "196x61+445+335"), ifelse(is_html_output(), "250%", "125%"))
project_team_img
```

Each project has team members assigned to it

* **Team_Member_Id:** Unique reference for each Team Member
* **Project_Id:** Id of Project

**Primary Key(s):** Composite. Team_Member_Id + Project_Id

**Foreign Key(s):** Team_Member_Id (Team_Member.Id), Project_Id (Project.Id). Each Project has multiple Team Members. Each team member can work on 1 or more projects.

```{r create project_team table, results='hide'}
dbExecute(conn, "CREATE TABLE Project_Team (
    Team_Member_Id integer NOT NULL,
    Project_Id integer NOT NULL,
    CONSTRAINT Project_Team_Id PRIMARY KEY (Project_Id,Team_Member_Id),
    CONSTRAINT Team_Team_Member FOREIGN KEY (Team_Member_Id)
    REFERENCES Team_Member (Id),
    CONSTRAINT Team_Project FOREIGN KEY (Project_Id)
    REFERENCES Project (Id)
);")
```

---

\newpage
### Role

```{r show Role table only, echo=FALSE}
role_img = image_scale(image_crop(er_img, "136x62+709+177"), ifelse(is_html_output(), "250%", "125%"))
role_img
```

Each team member has a role (and only 1 role as I already used 10 tables)

* **Id:** Unique reference for role
* **Name:** Role description

**Primary Key(s):** Id

**Foreign Key(s):** Id (Team_Member.Role_Id). Each team member has 1 role for projects.

```{r create role table, results='hide'}
dbExecute(conn, "CREATE TABLE Role (
    Id integer NOT NULL CONSTRAINT Role_pk PRIMARY KEY AUTOINCREMENT,
    Name varchar(50) NOT NULL
);")
```

---

\newpage
### Status

```{r show Status table only, echo=FALSE}
status_img = image_scale(image_crop(er_img, "138x63+708+34"), ifelse(is_html_output(), "250%", "125%"))
status_img
```

Work items have a progress status

* **Id:** Unique reference for status
* **Name:** Status description

**Primary Key(s):** Id

**Foreign Key(s):** Id (Work_Item.Stats_Id). Each work item has a status (To Do, In Progress, Review, Done).

```{r create Status table, results='hide'}
dbExecute(conn, "CREATE TABLE Status (
    Id integer NOT NULL CONSTRAINT Status_pk PRIMARY KEY AUTOINCREMENT,
    Name varchar(50) NOT NULL
);")
```

---

\newpage
### Team_Member

Describes an individual on the team

```{r show Team_Member table only, echo=FALSE}
team_member_img = image_scale(image_crop(er_img, "172x103+462+167"), ifelse(is_html_output(), "250%", "125%"))
team_member_img
```

* **Id:** Unique reference for team member
* **First_Name:** Team Member first name
* **Last_Name:** Team Member last name
* **Email:** Team Member email
* **Role_Id:** Team members role for projects

**Primary Key(s):** Id

**Foreign Key(s):** Id (Project_Team.Team_Member_Id, Work_Item.Assigned_To), Role_Id (Role.Id). Each Team Member has a role for projects. Each project team has team members. Each work item can be assigned to a team member.

```{r create team_member table, results='hide'}
dbExecute(conn, "CREATE TABLE Team_Member (
    Id integer NOT NULL CONSTRAINT Team_Member_pk PRIMARY KEY AUTOINCREMENT,
    First_Name varchar(50) NOT NULL,
    Last_Name varchar(50) NOT NULL,
    Email varchar(128) NOT NULL,
    Role_Id integer NOT NULL,
    CONSTRAINT Team_Member_Email_AK UNIQUE (Email),
    CONSTRAINT Team_Member_Role FOREIGN KEY (Role_Id)
    REFERENCES Role (Id)
);")
```

---

\newpage
### Work_Item

```{r show Work_Item table only, echo=FALSE}
work_item_img = image_scale(image_crop(er_img, "203x91+439+19"), ifelse(is_html_output(), "250%", "125%"))
work_item_img
```

Describes smallest parcel of work resulting in a testable output

* **Id:** Unique reference for work item.
* **Name:** Description of work item.
* **Status_Id:** Work item has as status (To Do, In Progress, Review, Done).
* **Assigned_To:** Work items can be assigned to team members.

**Primary Key(s):** Id

**Foreign Key(s):** Id (Asset_Work_Items.Work_Item.Id), Status_Id (Status.Id), Assigned_To (Team_Member.Id). Each work item can be assigned to a team member. Each work Item has a status. Assets are made of work items.

```{r create work_item table, results='hide'}
dbExecute(conn, "CREATE TABLE Work_Item (
    Id integer NOT NULL CONSTRAINT Work_Item_pk PRIMARY KEY AUTOINCREMENT,
    Name varchar(50) NOT NULL,
    Status_Id integer NOT NULL,
    Assigned_To integer NOT NULL,
    CONSTRAINT Work_Item_Statuses FOREIGN KEY (Status_Id)
    REFERENCES Status (Id),
    CONSTRAINT Work_Item_Team_Member FOREIGN KEY (Assigned_To)
    REFERENCES Team_Member (Id)
);")

```

---

\newpage
## Show tables in database

```{r list the tables}
dbListTables(conn)
```

---

## Data format function

Function to format tables for HTML and PDF output. Display tables using knitr library's kable function and kableExtra to format tables.

```{r}
data_format.function <- function(data, capt) {
  data %>%
    { 
      if (is_html_output()) { # if the output is HTML add class attribute
        kbl(., table.attr='class="table-striped table-hover"', caption = capt)} 
      else if (is_latex_output()) { # if the output is PDF ignore class attribute
        kbl(., caption = capt)
      }
    } %>% # pdf output keep tables in position
      kable_styling("striped", ifelse(is_html_output(),"hover","hold_position")) %>%
        row_spec(0, background = "#1AE81A")
}
```

---
---

\newpage
## Insert Data

### Status

```{r show fields in status table}
dbListFields(conn, "Status")
```

Insert table data (using DBI library dbExecute function):

```{r insert data into status table, results='hide'}
dbExecute(conn, "INSERT INTO Status ('Name') VALUES ('To Do');")
dbExecute(conn, "INSERT INTO Status ('Name') VALUES ('In Progress');")
dbExecute(conn, "INSERT INTO Status ('Name') VALUES ('Review');")
dbExecute(conn, "INSERT INTO Status ('Name') VALUES ('Done');")
```

Display table data (using knitr library kable function):

```{r show rows in status}
status_data <- dbGetQuery(conn, "SELECT * FROM Status;")
data_format.function(status_data, "Status Table")
```

\newpage
### Role

```{r show fields in role table}
dbListFields(conn, "Role")
```

Insert table data:

```{r insert data into role table, results='hide'}
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Project Manager');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Programmer');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Tester');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Artist');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('3D Modeller');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Environment Modeller');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Animator');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Shading Artist');")
dbExecute(conn, "INSERT INTO Role ('Name') VALUES ('Concept Artist');")
```

Display table data:

```{r show rows in role}
role_data <- dbGetQuery(conn, "SELECT * FROM Role;")
data_format.function(role_data, "Role Table")
```

\newpage
### Team_Member

```{r show fields in team_member table}
dbListFields(conn, "Team_Member")
```

Insert table data:

```{r insert data into team member table, results='hide'}
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Joe', 'O''Regan', 'joe.oregan@daie.ca4', 2);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Derp', 'McDerp', 'derpmcderp@daie.ca4', 1);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Herpderp', 'Derpderpenson', 'hd.derpderpenson@daie.ca4', 3);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Herpa', 'Derpderp', 'herpa.derpderp@daie.ca4', 4);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('De', 'Rpderp', 'de.rpderp@daie.ca4', 5);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Pred', 'Prehpred', 'predprehpred@daie.ca4', 6);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Derpa', 'Derpa', 'derpaderpa@daie.ca4', 9);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Herpa', 'Derpa', 'herpaderpa@daie.ca4', 8);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('HerpaDerpa', 'McDerpa', 'herpaderpa.mcderpa@daie.ca4', 7);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Joe', 'Derp', 'j.derp@daie.ca4', 2);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Jon', 'Herpaderp', 'jherpaderp@daie.ca4', 7);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Joblot', 'O''Stuff', 'joblot.ostuff@daie.ca4', 3);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Joderp', 'Herpderpenson', 'j.herpderpenson@daie.ca4', 4);")
dbExecute(conn, "INSERT INTO Team_Member ('First_Name', 'Last_Name', 'Email', 'Role_Id') 
          VALUES ('Jo', 'McQueryfiller', 'j.mcqueryfiller@daie.ca4', 1);")
```

Display table data:

```{r show rows in team member}
team_member_data <- dbGetQuery(conn, "SELECT * FROM Team_Member;")
data_format.function(team_member_data, "Team Member Table")
```

\newpage
### Work_Item

```{r show fields in Work_Item table}
dbListFields(conn, "Work_Item")
```

Insert table data:

```{r insert data into work_item table, results='hide'}
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Art Thingy', 2, 4);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Art Test Thingy', 3, 3);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Environment Model Thingy', 2, 6);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Art Concept Thingy', 4, 7);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Art Shading Thingy', 4, 8);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Random 3D Model', 2, 5);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('3D Model Test Obj', 2, 3);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Random Blueprint', 2, 1);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Blueprint Thingy', 2, 2);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Blueprint Test', 3, 2);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Art Test', 1, 12);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Query Model Thingy', 3, 11);")
dbExecute(conn, "INSERT INTO Work_Item ('Name', 'Status_Id', 'Assigned_To') 
          VALUES ('Another Test', 4, 13);")
```

Display table data:

```{r show rows in work_item}
work_item_data <- dbGetQuery(conn, "SELECT * FROM Work_Item;")
data_format.function(work_item_data, "Work Item Table")
```

\newpage
### Project

```{r show fields in Project table}
dbListFields(conn, "Project")
```

Insert table data:

```{r insert data into project table, results='hide'}
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Art Proj', '2023-01-11');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('DAIE CA4', '2023-01-20');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('New Project', '2023-01-24');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Old Project', '2022-12-14');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Christmas 2022 Project', '2022-12-25');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Date Range Project', '2023-01-17');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Another Date Range Project', '2023-02-01');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Project Filler', '2023-03-01');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Derp Project', '2023-03-17');")
dbExecute(conn, "INSERT INTO Project ('Name', 'Delivery_Date') 
          VALUES ('Hmmm I Ran Out of Names', '2023-01-20');")
```

Display table data:

```{r show rows in project}
project_data <- dbGetQuery(conn, "SELECT * FROM Project;")
data_format.function(project_data, "Project Table")
```

\newpage
### Project_Team

```{r show fields in Project_Team table}
dbListFields(conn, "Project_Team")
```

Insert table data:

```{r insert data into project_team table, results='hide'}
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (1, 1);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (2, 1);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (3, 1);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (4, 1);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (5, 1);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (2, 2);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (6, 2);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (7, 2);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (8, 2);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (9, 2);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (2, 3);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (10, 3);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (11, 3);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (14, 4);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (12, 4);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (13, 5);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (14, 6);")
dbExecute(conn, "INSERT INTO Project_Team ('Team_Member_Id', 'Project_Id') VALUES (14, 7);")
```

Display table data:

```{r show rows in project_team}
project_team_data <- dbGetQuery(conn, "SELECT * FROM Project_Team;")
data_format.function(project_team_data, "Project Team Table")
```

\newpage
### Asset

```{r show fields in Asset table}
dbListFields(conn, "Asset")
```

Insert table data:

```{r  insert data into asset table, results='hide'}
dbExecute(conn, "INSERT INTO Asset ('Name', 'Type', 'Format', 'Created_By', 'Date_Created') 
          VALUES ('Random Blueprint Asset', 'Combination of Blueprints', 'Zip file', 1, '2023-01-11');")
dbExecute(conn, "INSERT INTO Asset ('Name', 'Created_By', 'Date_Created') 
          VALUES ('Random Art Asset', 4, '2023-01-10');")
dbExecute(conn, "INSERT INTO Asset ('Name', 'Created_By', 'Date_Created') 
          VALUES ('Art Asset Thingy', 4, '2023-01-10');")
dbExecute(conn, "INSERT INTO Asset ('Name', 'Type', 'Created_By', 'Date_Created') 
          VALUES ('Environment Asset Thingy', 'Tree for use in Environment',4, '2023-01-02');")
```

Display table data:

```{r show rows in asset}
asset_data <- dbGetQuery(conn, "SELECT * FROM Asset;")
data_format.function(asset_data, "Asset Table")
```

\newpage
### Asset_Work_Items

```{r show fields in asset work items table}
dbListFields(conn, "Asset_Work_Items")
```

Insert table data:

```{r insert data into asset work items table, results='hide'}
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (8, 1);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (9, 1);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (10, 1);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (1, 2);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (2, 2);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (4, 3);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (5, 3);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (3, 4);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (6, 4);")
dbExecute(conn, "INSERT INTO Asset_Work_Items ('Work_Item_Id', 'Asset_Id') VALUES (7, 4);")
```

Display table data:

```{r show rows in Asset_Work_Items}
asset_work_items_data <- dbGetQuery(conn, "SELECT * FROM Asset_Work_Items;")
data_format.function(asset_work_items_data, "Asset Work Items Table")
```

\newpage
### Collection

```{r show fields in Collection table}
dbListFields(conn, "Collection")
```

Insert table data:

```{r insert data into collection table, results='hide'}
dbExecute(conn, "INSERT INTO Collection ('Library_Id', 'Asset_Id') VALUES (1, 1);")
dbExecute(conn, "INSERT INTO Collection ('Library_Id', 'Asset_Id') VALUES (2, 2);")
dbExecute(conn, "INSERT INTO Collection ('Library_Id', 'Asset_Id') VALUES (2, 3);")
dbExecute(conn, "INSERT INTO Collection ('Library_Id', 'Asset_Id') VALUES (3, 4);")
```

Display table data:

```{r show rows in collection}
collection_data <- dbGetQuery(conn, "SELECT * FROM Collection;")
data_format.function(collection_data, "Collection Table")
```

\newpage
### Library

```{r show fields in Library table}
dbListFields(conn, "Library")
```

Insert table data:

```{r insert data into library table, results='hide'}
dbExecute(conn, "INSERT INTO Library ('Name') VALUES ('Programming');")
dbExecute(conn, "INSERT INTO Library ('Name') VALUES ('Models');")
dbExecute(conn, "INSERT INTO Library ('Name') VALUES ('Scenery');")
dbExecute(conn, "INSERT INTO Library ('Name') VALUES ('Characters');")
```

Display table data:

```{r show rows in library}
library_data <- dbGetQuery(conn, "SELECT * FROM Library;")
data_format.function(library_data, "Library Table")
```

---

## Disconnect Database

```{r disconnect from database}
dbDisconnect(conn)
```

